---
title: "Social Participation Intersectionality code testing"
author: "Yan Liu and Tim Riffe"
date: "2024-01-08"
output: html_document
---

数据预处理，导入原始数据
Data cleaning, import original data
```{r}
library(haven)

df <- read_dta("E:/桌面/stata/社会参与/CLASSnew/class2020.dta")
```



政治参与变量 
D13近三年您是否参加过本地居民委员会/村民委员会的投票选举？
是... 1 
没参加，但是知道有选举活动 ... 2 
没参加，不知道是否有选举活动 ... 3

重新分组得到新变量policy
1：参加
0：未参加

political participation variable
D13：Have you participated in voting for local residents committee/village committee in the past three years?
yes: 1 
I didn’t participate, but I knew there was an election event: 2
I didn’t participate. I don’t know if there will be any election activities: 3

Regroup "D13" to get new variable "policy" 
1: Participate
0: Did not participate
```{r}
table(df$D13)
df$policy <- ifelse(df$D13 %in% c(1), 1, 0)
table(df$policy)
```


志愿活动变量 D15_3 和 D15_6
D15_3请问过去一年中，参加环境卫生保护的频率是？
D15_6.请问过去一年中，参加需要专业技术的志愿服务的频率是？ 
0:没参与
1：一年几次
2：每月至少一次
3：每周至少一次
4：几乎每天

重新分组D15_3 和 D15_6
1：参加
0：未参加

Volunteer activity variables D15_3 and D15_6
D15_3 In the past year, how often did you participate in environmental sanitation protection?
D15_6. In the past year, how often did you participate in volunteer services that required professional skills?
0: Not participating
1: Several times a year
2: At least once a month
3: At least once a week
4: Almost every day

Regroup "D15_3 and D15_6" to get new variable "nD15_3 and nD15_6" 
1: Participate
0: Not participating

```{r}
df$nD15_3 <- ifelse(df$D15_3 == 0, 0, 1)
df$nD15_6 <- ifelse(df$D15_6 == 0, 0, 1)
table(df$nD15_3)
table(df$nD15_6)
```

身体锻炼变量 G2
G2:您参加体育锻炼的频度是：
平均每月不足 1 次……………………………….1 
平均每月 1 次以上，但每周不足 1 次…2 
平均每周 1 次……………………………………….3 
平均每周 2 次……………………………………….4 
平均每周 3 次……………………………………….5 
平均每周 4 次……………………………………….6 
平均每周 5 次……………………………………….7 
平均每周 6 次……………………………………….8 
平均每周 7 次及以上…………………………….9 

重新分组D15_3 和 D15_6
(1/2=0) 0：每周不足 1 次以下
(3/9=1) 1：每周不足 1 次及以上

Physical exercise variable G2
G2: How often do you participate in physical exercise:
On average, less than 1 time per month……………………….1
On average, more than once a month, but less than once a week……………………….2
An average of 1 time per week…………………………………….3
Average 2 times a week…………………………………….4
An average of 3 times a week…………………………………….5
An average of 4 times a week…………………………………….6
Average 5 times per week…………………………………….7
Average 6 times per week…………………………………….8
Average of 7 times per week and above……………….9

Regroup Regroup "G2" to get new variable "exe" 
(1/2=0) 0: Less than once a week
(3/9=1) 1: Less than once a week or more
```{r}
table(df$G2)
df$exe <- ifelse(df$G2 %in% c(1, 2), 0, 1)
table(df$exe)
```


社会参与变量 D15_2_1 到 D15_2_6
过去一年中，您参加以下活动的情况是？
D15_2_1 宗教活动 
D15_2_2 上老年大学或者参加培训课程
D15_2_3 看电视/听广播/读书/看报/听戏
D15_2_4 唱歌/弹奏乐器 
D15_2_5 打麻将/下棋/打牌等
D15_2_6 广场舞

other Social participation variables D15_2_1 to D15_2_6
In the past year, how have you participated in the following activities?
D15_2_1 Religious activities
D15_2_2 Go to a senior college or participate in a training course
D15_2_3 Watching TV/listening to the radio/reading/reading newspapers
D15_2_4 Singing/playing an instrument
D15_2_5 Play mahjong/chess/cards, etc.
D15_2_6 Square Dance
Answer:
0: Not participating
1: Several times a year
2: At least once a month
3: At least once a week
4: Almost every day

Regroup "D15_2_1 ~ D15_2_6" to get new variable "nD15_2_1 ~ nD15_2_6" 
1: Participate
0: Not participating
```{r}

df$nD15_2_1 <- ifelse(df$D15_2_1 == 0, 0, 1)
df$nD15_2_2 <- ifelse(df$D15_2_2 == 0, 0, 1)
df$nD15_2_3 <- ifelse(df$D15_2_3 == 0, 0, 1)
df$nD15_2_4 <- ifelse(df$D15_2_4 == 0, 0, 1)
df$nD15_2_5 <- ifelse(df$D15_2_5 == 0, 0, 1)
df$nD15_2_6 <- ifelse(df$D15_2_6 == 0, 0, 1)

table(df$nD15_2_1)
table(df$nD15_2_2)
table(df$nD15_2_3)
table(df$nD15_2_4)
table(df$nD15_2_5)
table(df$nD15_2_6)
```

计算 sp 变量
Calculate the sum of sp
```{r}
df$sp <- df$policy + df$nD15_3 + df$nD15_6 + df$nD15_2_1 + df$nD15_2_2 + df$nD15_2_3 + df$nD15_2_4 + df$nD15_2_5 + df$nD15_2_6 + df$exe
table(df$sp)
```

 
A1.性别：
男 .. 1 
女 .. 2
A1. Gender:
1:Male  
2:Female  
Regroup "A1" to get new variable "male" 
1: male
0: female
```{r}
table(df$A1)
df$male <- ifelse(df$A1 == 1, "male", "female")
table(df$male)
```

收入变量 C10_1_1_open


For the measurement of personal income(C10_1_1_open), there are more than half of the missing values in the original data, so I used the variable with fewer missing values among the income-related variables that can be selected: household expenditure(C11_1_1_open) to perform linear interpolation on income.

1: <5000: Low income
2: 5001/10000: middle income
3: >10000: high income


收入变量 C10_1_1_open
Income variable C10_1_1_open
```{r}
df$C10_1_1_open[df$C10_1_1_open %in% c(999999,9999996,9999998, 9999999)] <- NA
sum(is.na(df$C10_1_1_open))
```

通过C11_1_1_open：“过去12个月，您家平均每月总支出多少” 线性插补
By C11_1_1_open: "What is your family's average monthly total expenditure in the past 12 months" Linear interpolation
```{r}
df$C11_1_1_open[df$C11_1_1_open %in% c(999998,999999)] <- NA
sum(is.na(df$C11_1_1_open))
```
删除用于插补的变量的缺失值
Remove missing values for variables used for imputation
```{r}
df <- subset(df, !is.na(C11_1_1_open))
```
进行线性插补
perform linear interpolation
```{r}
fit <- lm(C10_1_1_open ~ C11_1_1_open, data = df)
# 预测缺失值  Predict missing values
pred <- predict(fit, newdata = df[is.na(df$C10_1_1_open),])
# 替换缺失值为预测值 Replace missing values with predicted values
df$C10_1_1_open[is.na(df$C10_1_1_open)] <- pred 
```
分组编码
block coding
```{r}
df$income <- cut(df$C10_1_1_open, 
                 breaks = c(100, 5000, 10000, 408000),
                 labels = c("低收入", "中等收入", "高收入"))
table(df$income)
```


  

 
A10.您的户口属于？
农业户口 ............................................................... 1 
非农业户口 ........................................................... 2 
由农业户口改为统一居民户口 ........................... 3 
由非农业户口改为统一居民户口 ....................... 4 
其他（请注明：__________） .......................... 5
A10.Does your household registration belong to?
Agricultural household registration  ............... 1 
Non-agricultural household  ............ 2 
Change from agricultural household registration to unified residential household registration ............. 3
Change from non-agricultural household registration to unified resident household registration ............. 4
Others (please specify: __________)  ....... 5

group the "A10" to get new variable "hukou"
1：urban
0：not urban
```{r}
table(df$A10)
df$hukou <- ifelse(df$A10 == 1, "是", "否")
table(df$hukou)
```

A5_2 您的文化程度是：
不识字  ............... 1 
私塾/扫盲班 ................. 2 
小学  .................... 3 
初中  ................... 4 
高中/中专 ............... 5 
大专  .................. 6 
本科及以上 ……………………7 
A5_2 What is your education level:
Illiterate ............. 1
Private school/literacy class ............. 2
Elementary school ............. 3
Junior high school ............. 4
High School/Technical Secondary School............ 5
College ............. 6
Bachelor degree and above……………………7

group the "A5_2" to get new variable "edu"
0：1~2: 文盲
1: 3: 小学 
2: 4: 初中 
3: 5~7: 高中及以上 

0:1~2: Illiterate
1: 3: Elementary school
2: 4: junior high school
3: 5~7: High school and above

```{r}
table(df$A5_2)
library(dplyr)
df <- df %>%
  mutate(edu = case_when(
    A5_2 %in% c(1, 2) ~ 0,
    A5_2 == 3 ~ 1,
    A5_2 == 4 ~ 2,
    A5_2 %in% c(5, 6, 7) ~ 3,
    TRUE ~ NA_real_
  ))
levels(df$edu) <- c("文盲", "小学", "初中", "高中及以上")
attr(df$edu, "label") <- "受教育程度"
table(df$edu)
```

CES-D 抑郁量表:E2
1. 过去一周您觉得自己心情很好吗？  
2. 过去一周您觉得孤单吗？  
3. 过去一周您觉得心里很难过吗？  
4. 过去一周您觉得自己的日子过得很不错吗？ 
5. 过去一周您觉得不想吃东西吗？  
6. 过去一周您睡眠不好吗？ 
7. 过去一周您觉得自己不中用了吗？   
8. 过去一周您觉得自己没事可做吗？  
9. 过去一周您觉得生活中有很多乐趣（有意思的事情）吗？ 
答案：
1:没有 
2:有时 
3:经常 
9:无法回答
CES-D Depression Scale:E2
1. How good were you in your mood over the past week?
2. Have you felt lonely in the past week?
3. Have you felt sad in the past week?
4. How did you feel you had a good time in the past week?
5. Have you felt like you didn’t want to eat in the past week?
6. Did you have trouble sleeping in the past week?
7. Have you felt like you were useless this past week?
8. Did you feel like you had nothing to do in the past week?
9. Did you feel that there was a lot of fun (interesting things) in your life in the past week?
Answer:
1:No
2:sometimes
3: often
9: Unable to answer


首先将回答“9”定义为缺失值
然后，将E2_1 E2_4 E2_9三道题的答案转为 分数越高越抑郁 E2_1 E2_4 E2_9
最后将9项得分加总，以10分和17分为抑郁倾向和抑郁高风险的划界分
得到ced变量
0：无抑郁风险
1：抑郁倾向
2：抑郁高风险
First define the answer "9" as a missing value
Then, the answers to the three questions E2_1 E2_4 E2_9 are converted into The higher the score, the more depressed E2_1 E2_4 E2_9
Finally, the 9 scores were summed up, and 10 and 17 were used to delineate depression tendency and high risk of depression.
get “ced” variable
0: No risk of depression
1: Depressive tendencies
2: High risk of depression
```{r}
df[, c("E2_1", "E2_2", "E2_3", "E2_4", "E2_5", "E2_6", "E2_7", "E2_8", "E2_9")] <- lapply(df[, c("E2_1", "E2_2", "E2_3", "E2_4", "E2_5", "E2_6", "E2_7", "E2_8", "E2_9")], function(x) ifelse(x == 9, NA, x))

df$NE2_1 <- 4 - df$E2_1
df$NE2_4 <- 4 - df$E2_4
df$NE2_9 <- 4 - df$E2_9

df$ced1 <- df$NE2_1 + df$E2_2 + df$E2_3 + df$NE2_4 + df$E2_5 + df$E2_6 + df$E2_7 + df$E2_8 + df$NE2_9
df$ced <- cut(df$ced1, breaks = c(-Inf, 9, 16, 27), labels = c("0", "1", "2"), include.lowest = TRUE)
levels(df$ced) <- c("0", "1", "2")
table(df$ced)
```

自评健康
Self-rated health
B1. How do you feel about your current physical health?
Very healthy...................... 1
Relatively healthy ........................ 2
General........................3
Relatively unhealthy...................... 4
Very unhealthy ........................ 5
Unable to answer......................9

First define the answer "9" as a missing value
group the "B1" to get new variable "srh"
0:1~2:healthy
1:3:general
2:4~5:unhealthy
```{r}
df$B1[df$B1 == 9] <- NA
df$srh <- ifelse(df$B1 %in% c(1, 2), 0,
                 ifelse(df$B1 == 3, 1,
                        ifelse(df$B1 %in% c(4, 5), 2, NA)))
table(df$srh)
```

IADL
The instrumental daily living ability scale includes 10 indicators, namely, talking on the phone, dressing up, going up and down stairs, walking outside, taking public transportation, shopping, managing money, lifting, cooking, and doing housework. Each item is divided into three levels: "No help from others", "Need some help" and "Can't do it at all". If the answer is "can't do it at all" or "needs some help", it is judged to be disabled; if the answer is "no help from others", it is judged to be self-care.
* If 0 IADLs are disabled, the function is considered normal;
*Disablement of IADLs 1-5 is regarded as low function;
*Disability of 6-10 IADLs is regarded as significant functional impairment.

```{r}
df$nB4_1 <- ifelse(df$B4_1 %in% c(2, 3), 1, 0)
df$nB4_2 <- ifelse(df$B4_2 %in% c(2, 3), 1, 0)
df$nB6_1 <- ifelse(df$B6_1 %in% c(2, 3), 1, 0)
df$nB6_3 <- ifelse(df$B6_3 %in% c(2, 3), 1, 0)
df$nB6_4 <- ifelse(df$B6_4 %in% c(2, 3), 1, 0)
df$nB6_5 <- ifelse(df$B6_5 %in% c(2, 3), 1, 0)
df$nB6_6 <- ifelse(df$B6_6 %in% c(2, 3), 1, 0)
df$nB6_7 <- ifelse(df$B6_7 %in% c(2, 3), 1, 0)
df$nB6_8 <- ifelse(df$B6_8 %in% c(2, 3), 1, 0)
df$nB6_9 <- ifelse(df$B6_9 %in% c(2, 3), 1, 0)

df$IADL2020 <- df$nB4_1 + df$nB4_2 + df$nB6_1 + df$nB6_3 + df$nB6_4 + df$nB6_5 + df$nB6_6 + df$nB6_7 + df$nB6_8 + df$nB6_9
df$IADL <- cut(df$IADL2020, breaks = c(-Inf, 0, 5, 10), labels = c("0", "1", "2"), include.lowest = TRUE)
table(df$IADL)
```




查看缺失值
View missing values
```{r}
any(is.na(df$sp))
any(is.na(df$srh))
any(is.na(df$ced))
any(is.na(df$IADL))
any(is.na(df$edu))
any(is.na(df$hukou))
any(is.na(df$income))
```
删除缺失值
Remove missing values
```{r}
df <- subset(df, !is.na(srh))
df <- subset(df, !is.na(ced))
df <- subset(df, !is.na(edu))
df <- subset(df, !is.na(income))
```

View all variables used
Y：
SES:edu、hukou、income
health condition：srh、ced、IADL
male
```{r}
table(df$sp)
table(df$srh)
table(df$ced)
table(df$IADL)
table(df$edu)
table(df$hukou)
table(df$income)
table(df$male)
```

Keep the variables used in our next analysis to Df
```{r}
Df <- df[, c("sp", "srh", "ced", "IADL", "edu", "hukou", "income", "male")]

```
 

#LCA analysis needs to be replaced 


LCA analysis
```{r}
library(poLCA)
library(haven)
Data <- read_dta("E:/桌面/R/新建文件夹/myMAIHDA/11.dta")
```


```{r, eval = FALSE}
vars <- c("ced", "srh", "IADL", "edu","hukou")
for (i in 1:length(vars)){
	Data[,vars[i]] <- Data[,vars[i]]+1
}
```


```{r}
f <- as.formula(cbind(cedd, srhh, IADLL)~1)
LCA2 <- poLCA(f, data=Data, nclass=2)
```


```{r}
#slide13:
plot(LCA2)
#predicted class membership is in:
LCA2$predclass[1:30]
#could be used as another variable (part of the data):
Data$health <- LCA2$predclass
```


```{r}
h<- as.formula(cbind(incomenew, edu, hukuo)~1)
lca2 <- poLCA(f, data=Data, nclass=2)
#slide13:
plot(lca2)
#predicted class membership is in:
lca2$predclass[1:30]
#could be used as another variable (part of the data):
Data$ses <-lca2$predclass
```





MAIHDA analysis
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, eval = FALSE}
install.packages("brms")
install.packages("ggmcmc")
install.packages("arm")
install.packages("survey")
install.packages("srvyr")
install.packages("jtools")
install.packages("cowplot")
install.packages("expss")
install.packages("ggh4x")
install.packages("haven")
install.packages("miceadds")
install.packages("rio")
install.packages("tidyverse")
install.packages("WriteXLS")
install.packages("parallel")
install.packages("future")
install.packages("tictoc")
install.packages("tidybayes")
install.packages("readxl")
```

```{r, message = FALSE, warning=FALSE}
library(brms)
library(ggmcmc)
library(arm)
library(survey)
library(srvyr)
library(jtools)
library(cowplot)
library(expss)
library(ggh4x)
library(haven)
library(miceadds)
library(rio)
library(tidyverse)
library(WriteXLS)
library(parallel)
library(future)
library(tictoc)
library(tidybayes)
library(readxl)

set.seed(195)

```


Read data 
```{r}
data <- haven::read_dta("Data/new-data.dta")

head(data)
#列出名为data的对象中所包含的所有变量或对象的名称列表
ls(data)

#创建一个频数表/频率表来统计data数据集中male变量的不同取值及其对应的出现次数
#原文的应用是虚拟变量的情况
table(data$male)

```

为了确保该变量被正确解释为数值型，在这里我们首先使用 as.character() 将其转换为字符型，然后再使用 as.numeric() 将其转换为数值型，并将结果重新赋值给 data
To ensure that the variable is correctly interpreted as a numeric value,Here we first use as.character() to convert it to a character type,Then use as.numeric() to convert it to a numeric type,and reassign the result to data
```{r}
data$male  <- as.numeric(as.character(data$male))
data$health  <- as.numeric(as.character(data$health))
data$ses  <- as.numeric(as.character(data$ses))
data$sp  <- as.numeric(as.character(data$sp))
```

对数据框中的变量进行频数统计，并输出一个频数表。这将帮助你了解变量的分布情况
performs frequency statistics on the variables in the data frame and outputs a frequency table.To understand the distribution of the variable
```{r}
table(data$sp)
table(data$male)
table(data$health)
table(data$ses)
```

查看变量的数据结构
View variable data structure
```{r}
str(data$sp)
str(data$male)
str(data$health)
str(data$ses)

str(data$id)
# rid is ID ,like the country,抓换成character

as.character(data$id)

str(data$id)

```

将名为 data的对象转换为一个数据框，并将结果重新赋值给 data
Convert the object named data into a data frame,and reassign the result to data
```{r}
data <- as.data.frame(data)
```

将处理后的结果重新赋值给data数据框的所有列，
从而移除了每列的属性。
通常用于简化数据的结构，
特别是在数据预处理阶段。移除属性后，
数据框中的每个变量只保留其基本的数据结构，
不再包含额外的属性信息。
这有助于确保数据的干净和简洁，使其更适合用于后续的分析和建模。
请注意，这行代码将修改data数据框本身，
并且在运行后，data的所有列将不再包含额外的属性。
Reassign the processed results to all columns of the data frame,
Thus removing the attribute for each column.
Often used to simplify the structure of data,
Especially in the data preprocessing stage. After removing the attribute,
Each variable in the data frame retains only its basic data structure,
No additional attribute information is included.
This helps ensure that the data is clean and concise, making it more suitable for subsequent analysis and modeling.
Note that this line of code will modify the data frame itself,
And after running, all columns of data will no longer contain extra attributes.

```{r}
data[] <- lapply(data, function(x) {attributes(x) <- NULL;x})
```

查看变量的数据结构
View variable data structure
```{r}
str(data$sp)
str(data$male)
str(data$health)
str(data$ses)
str(data$id)
```


你将得到一个名为 std 的列表，
其中包含所有变量的均值和标准差信息。
命令经过修改：将data[,v]改为了data[[v]]
这是因为你的数据框data中的列名是字符向量
所以在循环中使用双括号[[ ]]来提取对应的列数据。
现在，这个循环应该能够计算"sp"、"male"、
"ses"和"health"这四个列的均值和标准差，并将结果存储在std列表中。
You will get a list named std,
It contains mean and standard deviation information for all variables.
The command has been modified: changed data[,v] to data[[v]]
This is because the column names in your data frame data are character vectors
So use double brackets [[ ]] in the loop to extract the corresponding column data.
Now, the loop should be able to calculate "sp", "male",
The mean and standard deviation of the four columns "ses" and "health" and store the result in the std list.

```{r}

new_data <-
  data |> 
  group_by(male, health, ses) |> 
  mutate(strata = cur_group_id()) |> 
  select(id,strata,male,health,ses,sp) |> 
  pivot_longer(-c(id,strata), names_to = "variable", values_to = "value") |> 
  group_by(variable) |> 
  mutate(n_value = (value - mean(value)) / sd(value)) |> 
  ungroup() |> 
  select(-value) |> 
  pivot_wider(names_from = variable, values_from = n_value) |> 
  ungroup() 
```
 





# ---- MAIHDA -----

Add strata
2 (male) x 2 (health) x 2 (ses) = 8 strata 


Form intersectional strata Check whether sample sizes in strata are sufficiently large Assign a strata ID to each individual in the data set
步骤 1形成交叉分层检查分层中的样本量是否足够大为数据集中的每个个体分配分层 ID

```{r}
#Example: Calculation of the percentage of strata with 10 or more individuals
#Store total number of strata in an object:

strata_n <- new_data |> 
  group_by(strata) |> 
  summarize(n = n())

```





Perform multilevel analysis to partition the  variance between and  within intersectional  strata 
执行多级分析以划分交叉层之间和内部的方差

(1) Variance decomposition in the simple intersectional model: Total variance in the outcome is decomposed into variance that can be attributed to
(a) mean-level differences between intersectional strata and 
(b) interindividual differences within intersectional strata

(1) 简单交叉模型中的方差分解：结果的总方差被分解为可归因于 
(a) 交叉层之间的平均水平差异和
(b) 交叉层内个体间差异的方差


```{r, eval = FALSE}
#Calculate simple intersectional model1:

model1 <- brm(sp ~ 1 + (1|strata),
              data = new_data,
              warmup = 5000, 
              iter = 10000, 
              chains = 1, seed = 123)


#Extract variance:
model1


```




(2) Variance decomposition in the intersectional interaction model: 
    The intersectional interaction model extends the simple intersectional model by including the  additive main effects of different social categories at the strata level (as fixed effects) to compute the between- and within-strata variance after accounting for main effects
(2) 交叉交互模型中的方差分解：
   交叉交互模型扩展了简单交叉交互模型模型通过包括阶层层面不同社会类别的加性主效应（作为固定效应）来计算考虑主效应后的阶层间和阶层内方差

```{r, eval = FALSE}
#Calculate intersectional interaction model1

model2 <- brm(sp ~ 1 + male + ses + health + (1|strata),
              data = new_data,
              warmup = 5000, 
              iter = 10000, 
              chains = 1, seed = 123)

#Extract variance:
model2
pairs(model2)
```
 
Read in the previously-fit model results:
```{r}
model1 <- readRDS("Data/new-MAIHDA_model1.rds")
model2 <- readRDS("Data/new-MAIHDA_model2.rds")
```



Data analysis: Analyses to answer the research questions
数据分析：分析回答研究问题
 

To what extent do complex interactions of social categories contribute incrementally to explaining sp inequalities?
(1) Calculate the strata-level varianceattributable only to interaction effects, at least with respect to the variables included, by quantifying the adjusted VPC in the intersectional interaction model
RQ 3：社会类别的复杂相互作用在多大程度上有助于解释社会参与不平等？
(1) 通过量化交叉交互模型中调整后的VPC，计算仅归因于交互效应的层级方差，至少就所包含的变量而言


```{r}
# Variance at strata level model 1 分层水平的方差是v1=0.07435116
coef_mat1 <- 
  model1 |> 
  posterior_summary() |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column("coef") 
v1 <-
  coef_mat1 |> 
  filter(coef == "sd_strata__Intercept") |> 
  pull("Estimate") %>%
  '^'(2)
# Variance at individual level model 1 个体水平的方差s1=0.957085
s1 <-
  coef_mat1 |> 
  filter(coef == "sigma") |> 
  pull("Estimate") %>%
  '^'(2)
  
# Calculate VPC model 1  #7.21
round(100 * v1 / (v1 + s1),2)
 

```

```{r}
# Variance at strata level model 2分层水平的方差是 v2=0.002062565
coef_mat2 <- 
  model2 |> 
  posterior_summary() |> 
  as.matrix() |> 
  as.data.frame() |> 
  rownames_to_column("coef") 
v2 <-
  coef_mat2 |> 
  filter(coef == "sd_strata__Intercept") |> 
  pull("Estimate") %>%
  '^'(2)
# Variance at individual level model 2 个体水平的方差s2=0.9573226
s2 <-
  coef_mat2 |> 
  filter(coef == "sigma") |> 
  pull("Estimate") %>%
  '^'(2)

 # Calculate VPC model 2  =0.21
round(100 * v2 / (v2 + s2),2)
 
 
```
 
(2) Calculate the percentage of between-strata variance explained by the main effects by quantifying the PCV. The difference from 100% corresponds to the proportion of between-strata variance that is (likely) due to the existence of interaction effects between the social categories defining the strata(2)通过量化PCV，计算由主效应解释的层间方差的百分比。 与 100% 的差异对应于阶层间方差的比例，这（可能）是由于定义阶层的社会类别之间存在相互作用效应
 
```{r}
#Calculate the PCV (in %): =2.774085
100 - 100 * (v1 - v2) / v1
#Calculate the percentage of between-strata variance that cannot be explained by main effects (in %):

# 100-pcv
 
```
 
 

  

Extract random effects、 predicted and expected values
```{r}

#### Extract random effects
bayes.random.effects <- brms::ranef(model2)

# As data frame
bayes.random.effects.new <- as.data.frame(bayes.random.effects$strata)

#Round
bayes.random.effects.new <- round(bayes.random.effects.new, 3)

# Add strata number
bayes.random.effects.new$strata <- 1:8

# Change order of variables
bayes.random.effects.new <- bayes.random.effects.new[,c(5,1,2,3,4)]



strata_def <- 
  data |> 
  group_by(male, health, ses) |> 
  summarize(strata = cur_group_id(), .groups = "drop") 
bayes.random.effects.new |> colnames()
bayes.random.effects.new |> 
  left_join(strata_def, by= "strata") |> 
  mutate(sex = if_else(male == 0, "female","male"),
         health = if_else(health == 1, "healthy", "unhealthy"),
         ses = if_else(ses == 1, "high ses", "low ses")) |> 
  ggplot(aes(y = reorder(strata,Estimate.Intercept), x = Estimate.Intercept)) +
  geom_point() +
  theme_minimal() +
  geom_pointrange(aes(xmin = Q2.5.Intercept, xmax = Q97.5.Intercept)) +
  geom_vline(xintercept = 0) +
  labs(title = "Random effects",
       subtitle = "Interpreted as interractions (?)",
       caption = "strata sorted by point estimate value",
       y = "strata")

```

```{r}
#### Extract predicted and expected values

## This analysis is computationally very extensive. It might be sensible to calculate the predicted values and save 
## the results, restart the script to conduct the rest of the analyses and load the results again when needed to 
## plot the data.


# Predicted average social participation per strata
 
predicted.means.model2 <-
                 model2 %>%
                 epred_draws(newdata =new_data) %>% 
                 group_by(strata) %>%
                 mean_qi(.epred)

predicted.means.model2 |> 
  left_join(strata_def, by= "strata") |> 
  mutate(sex = if_else(male == 0, "female","male"),
         health = if_else(health == 1, "healthy", "unhealthy"),
         ses = if_else(ses == 1, "high ses", "low ses")) |> 
  ggplot(aes(y = reorder(strata,.epred), x = .epred)) +
  geom_point() +
  theme_minimal() +
  geom_pointrange(aes(xmin = .lower, xmax = .upper)) +
  geom_vline(xintercept = 0) +
  labs(title = "predicted strata sp means",
       subtitle = "comparable with empirical means",
       caption = "strata sorted by point estimate value",
       y = "strata",
       x = "strata mean sp values")

# save(predicted.means.model1, file = "predicted.means.model1.rds")
#load("C:/Users/File/predicted.means.model1.rds")
```


```{r}

# Expected average social participation per strata

expected.means.model2 <- model2 %>%
                             epred_draws(newdata = new_data, re_formula = NA) %>%
                             group_by(strata) %>%
                             mean_qi(.epred)

View(expected.means.model2)
expected.means.model2 |> 
  left_join(strata_def, by= "strata") |> 
  mutate(sex = if_else(male == 0, "female","male"),
         health = if_else(health == 1, "healthy", "unhealthy"),
         ses = if_else(ses == 1, "high ses", "low ses")) |> 
  ggplot(aes(y = reorder(strata,.epred), x = .epred)) +
  geom_point() +
  theme_minimal() +
  geom_pointrange(aes(xmin = .lower, xmax = .upper)) +
  geom_vline(xintercept = 0) +
  labs(title = "expected strata sp means",
       subtitle = "comparable with empirical means",
       caption = "strata sorted by point estimate value",
       y = "strata",
       x = "strata mean sp values")
```

```{r}
# save(expected.means.model1, file = "expected.means.model1.rds")
#load("C:/Users/File/expected.means.model2.rds")




# Check whether difference between predicted and expected values equals random effects
# show2 <- tibble(pred.mean2 = predicted.means.model2$sp, 
                # expect.mean2 = expected.means.model2$sp,
                # ranef_int2 = bayes.random.effects.new$Estimate.Intercept)
# show2 |> 
  # mutate(pred.mean2 - expect.mean2)

# show2$diff <- show2$V1 - show2$V2
# View(show2) # difference equals random effects



 
# Check whether difference between predicted and expected values equals random effects
show2 <- cbind(predicted.means.model2$.epred, expected.means.model2$.epred, bayes.random.effects.new$Estimate.Intercept)
show2 <- as.data.frame(show2)
show2$diff <- show2$V1 - show2$V2
View(show2) # difference equals random effects

# TR: predicted mean appears to be the same as the expected marginal mean from the model, plus the
# random effects from the model. So the expected means are the pure margins of the model I guess
# I'm just trying to reverse engineer what everything means.

# TR: these are the same! So something isn't right 
# LY:According to this code, there is a difference
 
```

Stop here:
 
 
Merge average expected and predicted reading achievement per stratum
```{r}
predicted.means.model1 <- predicted.means.model1[,1:4]
colnames(predicted.means.model1) <- c("strata", "predicted.mean.model1", "predicted.mean.model1.CI.lb", "predicted.mean.model1.CI.ub")

predicted.means.model2 <- predicted.means.model2[,1:4]
colnames(predicted.means.model2) <- c("strata", "predicted.mean.model2", "predicted.mean.model2.CI.lb", "predicted.mean.model2.CI.ub")

expected.means.model2 <- expected.means.model2[,1:4]
colnames(expected.means.model2) <- c("strata", "expected.mean.model2", "expected.mean.model2.CI.lb", "expected.mean.model2.CI.ub")

predicted.expected.means <- full_join(predicted.means.model1, expected.means.model2, by = "strata") %>%             
  full_join(., predicted.means.model2, by = "strata") 
View(predicted.expected.means)
```


 
   





```{r}

# ---- Figure 2: Predicted values in strata x SP levels ---- #


# Merge data with predicted values per stratum (model 1 and 2)
df <- new_data 
data.plot.new <- merge(df, predicted.means.model1, by = "strata")
data.plot.new <- merge(data.plot.new, predicted.means.model2, by = "strata")


# Rename values for plotting
# LY: I guess we don't need to "Rename values for plotting"
#data.plot.new <- data.plot.new %>% 
  #dplyr::mutate(male = dplyr::recode(male, 
                                     #  `0`="female",
                                     #  `1`="male"))


#data.plot.new <- data.plot.new %>% 
 # dplyr::mutate(health = dplyr::recode(health, 
                                            #  `1`="health",
                                             # `2`="unhealth"))

#data.plot.new <- data.plot.new %>% 
  #dplyr::mutate(ses = dplyr::recode(ses, 
                                             #   `1`="high ses",
                                             #   `2`="low ses"))

 


# Subset data
data.plot.model1 <- subset(data.plot.new, select = c("strata", "male", "ses", "health", 
                                                      "predicted.mean.model1", 
                                                     "predicted.mean.model1.CI.lb", "predicted.mean.model1.CI.ub"))


# Reduce data
data.plot.model1 <- unique(data.plot.model1)




# Plot
figure2 <- ggplot(data = data.plot.model1, aes(x = predicted.mean.model1, y = gender, color = gender))
figure2 <- figure2 + geom_rect(aes(xmin = 334.75, xmax = 407.47, ymin = -0.5, ymax = Inf, fill = "grey60"), colour = NA, alpha = 0.3)
figure2 <- figure2 + geom_rect(aes(xmin = 407.47, xmax = 480.18, ymin = -0.5, ymax = Inf, fill = "grey70"), colour = NA, alpha = 0.3)
figure2 <- figure2 + geom_rect(aes(xmin = 480.18, xmax = 552.89, ymin = -0.5, ymax = Inf, fill = "grey80"), colour = NA, alpha = 0.3)
figure2 <- figure2 + geom_rect(aes(xmin = 552.89, xmax = 625.61, ymin = -0.5, ymax = Inf, fill = "grey90"), colour = NA, alpha = 0.3)
figure2 <- figure2 + geom_errorbar(aes(xmin = predicted.mean.model1.CI.lb, xmax = predicted.mean.model1.CI.ub))
figure2 <- figure2 + geom_point()
figure2 <- figure2 + facet_nested(~ male+ health + ses)
figure2 <- figure2 + coord_flip(clip="off")
figure2 <- figure2 + theme_classic() 
figure2 <- figure2 + xlab("Predicted mean sp") + ylab("")
figure2 <- figure2 + theme(axis.text.x = element_text(size = 11.5, angle = 90, vjust = 0.5, hjust=1),
                           axis.text.y = element_text(size = 11.5),
                           axis.title.y = element_text(size = 14),
                           strip.placement = "inside",
                           strip.text = element_text(size = 11.5),
                           legend.position = "bottom",
                           legend.text = element_text(size = 11.5),
                           legend.title = element_text(size = 13))
figure2 <- figure2 + scale_x_continuous(breaks=seq(325, 625, 25))
figure2 <- figure2 + scale_y_continuous(labels=c("Male","Female"), breaks=c(1,0))
figure2 <- figure2 + scale_color_gradient(low = "firebrick3", high ="#083563")
figure2 <- figure2 + scale_fill_manual('PISA proficiency level', values = c("grey60", "grey70", "grey80","grey90"),
                                       labels = c("1a", "2", "3", "4"),
                                       guide = guide_legend(override.aes = list(alpha = 0.3)))
figure2 <- figure2 + scale_shape_manual(values = c(16, 17))
figure2 <- figure2 + guides(color = "none")
figure2



# Set working directory (e.g., "C:/users/Your.File")
setwd("")

png("figure2.png", width = 16, height = 6, units = "in", res=300)
plot_grid(figure2)
dev.off()






# ---- Figure 3: Plot random effects ---- #

# Select first imputed data frame
df <- new_data 

# Merge data
data <- merge(bayes.random.effects.new, df, by = "strata")


# Subset relevant variables for table and plotting
data.plot <- subset(data, select = c("strata", "male", "health", "ses", 
                                     "sp"))
data.plot <- unique(data.plot)


# Rename values for plotting
data.plot <- data.plot %>% 
  dplyr::mutate(male = dplyr::recode(male, 
                                       `0`="female",
                                       `1`="male"))

data.plot <- data.plot %>% 
  dplyr::mutate(health = dplyr::recode(health, 
                                              `1`="health",
                                              `2`="unhealth"))

data.plot <- data.plot %>% 
  dplyr::mutate(HISCED.abitur = dplyr::recode(HISCED.abitur, 
                                              `0`="Below university entrance certificate",
                                              `1`="At least university entrance certificate"))

data.plot <- data.plot %>% 
  dplyr::mutate(ses = dplyr::recode(ses, 
                                                `1`="high ses",
                                                `2`="low ses"))

 


# Subset data
data.plot.model1 <- subset(data.plot.new, select = c("strata", "male", "ses", "health", 
                                                      "predicted.mean.model1", 
                                                     "predicted.mean.model1.CI.lb", "predicted.mean.model1.CI.ub"))






 


### Plot figure 2
figure3 <- ggplot(data = data.plot, aes(x = Estimate.Intercept, y = gender, color = gender))
figure3 <- figure3 + geom_errorbar(aes(xmin = Q2.5.Intercept, xmax = Q97.5.Intercept))
figure3 <- figure3 + geom_point() 
figure3 <- figure3 + facet_nested(~  male+ health + ses)
figure3 <- figure3 + coord_flip(clip="off", xlim = c(-45, 45))
figure3 <- figure3 + theme_classic() 
figure3 <- figure3 + xlab(expression(mu[0][j])) + ylab("")
figure3 <- figure3 + theme(axis.text.x = element_text(size = 11.5, angle = 90, vjust = 0.5, hjust=1),
                           axis.text.y = element_text(size = 11.5),
                           axis.title.y = element_text(size = 14),
                           strip.placement = "inside",
                           strip.text = element_text( size = 11.5),
                           legend.position = "none")
figure3 <- figure3 + scale_x_continuous(breaks=seq(-50, 50, 5))
figure3 <- figure3 + scale_color_manual(values=c("#083563", "firebrick3"))
figure3



# Set working directory (e.g., "C:/users/Your.File")
setwd("")

png("figure3.png", width = 15, height = 6, units = "in", res=300)
plot_grid(figure3)
grid.lines(x=c(0.046,0.995), y = c(0.497,0.497), gp = gpar(col = "grey40", lty=2))
dev.off()


# Figure 3 with arrows above
png("figure3.png", width = 15, height = 6, units = "in", res=300)
plot_grid(figure3)
grid.lines(x=c(0.046,0.995), y = c(0.497,0.497), gp = gpar(col = "grey40", lty=2))
grid.lines(x = unit(c(0.9835, 0.9835), "npc"), y = unit(c(0.835, 0.788), "npc"), gp = gpar(fill="black"),
           arrow = arrow(length = unit(0.11, "inches"), ends="last", type="closed"))
dev.off()



```







